Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
setwd("C:/Umesh JN/Personal/Learning/Data Science/NanoDegree-Udacity/DataAnalysiswithR/code")
pf <- read.delim("pseudo_facebook.tsv")

library(ggplot2)
ggplot(aes(x = gender, y = age),data = subset(pf, !is.na(gender))) + 
  geom_boxplot() +
  stat_summary(fun.y=mean,geom="point",shape=4)


ggplot(aes(x = age, y = friend_count),data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color=gender),stat="summary",fun.y=median)


library(dplyr)

pf <- filter(pf,!is.na(gender))
pf_age_gender <- group_by(pf,age,gender)
pf.fc_by_age_gender <- summarise(pf_age_gender,mean_friend_count = mean(friend_count),median_friend_count = median(as.numeric(friend_count)),n=n())
pf.fc_by_age_gender <- arrange(pf.fc_by_age_gender,age)

head(pf.fc_by_age_gender)

#Using chaining syntax
fc_by_age_gender_chaining <- pf %>% filter(!is.na(gender)) %>% group_by(age,gender) %>% summarise(mean_friend_count = mean(friend_count),median_friend_count=median(as.numeric(friend_count)),n=n()) %>% arrange(age)


head(fc_by_age_gender_chaining)

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),data = pf.fc_by_age_gender) + geom_line(aes(color=gender))

ggplot(aes(x = age, y = median_friend_count),data = pf.fc_by_age_gender) + geom_point(aes(color=gender)) + geom_line(aes(color=gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('tidyr')
install.packages('dplyr')
library(tidyr)
library(dplyr)
head(pf.fc_by_age_gender)
pf.select_age_gender_median <- select(pf.fc_by_age_gender,age,gender,median_friend_count)
pf.reshaped_age_gender_median <- spread(pf.select_age_gender_median,gender,median_friend_count)

pf.fc_by_age_gender.wide <- pf.fc_by_age_gender %>% select(age,gender,median_friend_count) %>% spread(gender,median_friend_count)

head(pf.reshaped_age_gender_median)
head(pf.fc_by_age_gender.wide)

```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
pf.fc_by_age_gender.wide$ratio <- pf.fc_by_age_gender.wide$female / pf.fc_by_age_gender.wide$male
head(pf.fc_by_age_gender.wide)

ggplot(aes(x = age, y = ratio),data = pf.fc_by_age_gender.wide) + geom_point() +geom_line()  + geom_hline(linetype=2,yintercept=1)

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
head(pf$tenure,10)
pf$year_joined <- floor(2014 - pf$tenure/365)

head(pf$year_joined,10)

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket <- cut(pf$year_joined,breaks=c(2004,2009,2011,2012,2014))

table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
p1 <- ggplot(aes(x = age, y = friend_count),data = pf) + geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=median)


p2 <- ggplot(aes(x = age, y = friend_count),data = subset(pf,!is.na(year_joined.bucket))) + geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=median)


library(gridExtra)

grid.arrange(p1,p2)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),data = subset(pf,!is.na(year_joined.bucket))) + geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=mean) + geom_line(stat="summary",fun.y=mean,linetype=3)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
head(pf)
pf_rate <- subset(pf,tenure >=1 )
head(pf_rate)
# pf_rate <- subset(pf, tenure >= 1 && friend_count >= 1)
pf_rate$rate <- pf_rate$friend_count/pf_rate$tenure

head(pf_rate$rate)
tail(pf_rate$rate)

summary(pf_rate$rate)

with(subset(pf,tenure>=1),summary(friend_count/tenure))

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
pf_friendship_initiated <- subset(pf,tenure >=1 )

pf_friendship_initiated$fs_initiation_rate <- pf_rate$friendships_initiated/pf_rate$tenure

ggplot(aes(x=tenure,y=fs_initiation_rate),data=pf_friendship_initiated) + 
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=mean)


ggplot(aes(x=tenure,y=friendships_initiated/tenure),data=subset(pf,tenure >=1 )) + 
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)


ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color=year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

yo <- read.csv("yogurt.csv")
head(yo)
tail(yo)
str(yo)
yo$id <- factor(yo$id)


ggplot(yo,aes(price)) +
  geom_histogram(binwidth=1) +
  scale_x_continuous(breaks=seq(0,70,5))


ggplot(yo,aes(price)) +
  geom_histogram(binwidth=5,fill="blue",color="red") +
  scale_x_continuous(breaks=seq(0,70,5)) + 

summary(yo$price)

yo$all.purchases <- sum(c("strawberry", "blueberry","pina.colada","plain","mixed.berry"))

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo <- transform(yo,all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
table(yo$all.purchases)
summary(yo$all.purchases)

ggplot(yo,aes(all.purchases)) + 
  geom_histogram(binwidth=1)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
head(yo,2)
tail(yo,2)

ggplot(yo,aes(x=time,y=price)) +
  geom_jitter(alpha=1/20)

ggsave("price_vs_time.png",dpi=300,width=10,height=8)
ggsave("price_vs_time.png",dpi=300,width=10,height=8)


ggplot(data = subset(yo,c(strawberry,blueberry,pina.colada,plain,mixed.berry))) + geom_bar()
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

set.seed(4230)
sample.ids <- sample(levels(yo$id),16)

ggplot(aes(x=time,y=price),data=subset(yo,id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size=all.purchases),pch=1)


set.seed(21)
sample.ids <- sample(levels(yo$id),16)

ggplot(aes(x=time,y=price),data=subset(yo,id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size=all.purchases),pch=1)



```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
install.packages('GGally')
library(GG)

```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

