Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
setwd("C:/Umesh JN/Personal/Learning/Data Science/NanoDegree-Udacity/DataAnalysiswithR/code")
pf <- read.delim("pseudo_facebook.tsv")

library(ggplot2)
qplot(x=age,y=friend_count,data=pf)
qplot(age,friend_count,data=pf)

```

***

#### What are some things that you notice right away?
Response:
Majority of the users are within the age group below 30. 

Also these young people below the age of 30  have friends count more than few thousands and some even having around 5K

There seems to be some users who are older than 90 but those seems to be fake ones as it is very unlikely for people so old to be having so many friends and being active in FB.
***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x=age,y=friend_count),data=pf) + geom_point()

summary(pf$age)

ggplot(aes(x=age,y=friend_count),data=pf) + geom_point() + xlim(13,90)

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x=age,y=friend_count),data=pf) + geom_point(alpha=1/20) + xlim(13,90) 

ggplot(aes(x=age,y=friend_count),data=pf) + geom_jitter(alpha=1/20) + xlim(13,90) 

```

#### What do you notice in the plot?
Response:
Majority of the users have less than 350 or 400 friends.

Also users less than the age of 25 have more friends compared to other age groups.

There are very few users having more than thousands users and also who fall in the age group above 75.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age,y=friend_count),data=pf) + geom_point(alpha=1/20) + xlim(13,90) +   coord_trans(y="sqrt")

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x=age,y=friend_count),data=pf) + geom_point(alpha=1/20, position = position_jitter(h=0)) + xlim(13,90) +   coord_trans(y="sqrt")
```

#### What do you notice?
Lot of users have zero friends.
Also majority of them have the friends below the 500.
***

### Alpha and Jitter
Notes:

Younger users below the age of 25 tend to initiate more friendship requests compared to other age groups.

There seems to be completely different trend only for the age 69 again might be due to fake accounts by the active user groups.


```{r Alpha and Jitter}
names(pf)

ggplot(aes(x=age,y=friendships_initiated),data=pf) + geom_point()

ggplot(aes(x=age,y=friendships_initiated),data=pf) + geom_point() + xlim(13,90)

ggplot(aes(x=age,y=friendships_initiated),data=pf) + geom_point(alpha=1/20) + xlim(13,90)

ggplot(aes(x=age,y=friendships_initiated),data=pf) + geom_jitter(alpha=1/20) + xlim(13,90)

ggplot(aes(x=age,y=friendships_initiated),data=pf) + geom_point(alpha=1/20) + xlim(13,90) + coord_trans(y="sqrt")

ggplot(aes(x=age,y=friendships_initiated),data=pf) + geom_point(alpha=1/20,position=position_jitter(h=0)) + xlim(13,90) + coord_trans(y="sqrt")

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

age_group <- group_by(pf,age)
pf.fc_by_age <- summarise(age_group,fc_mean = mean(friend_count),fc_median=median(friend_count), n = n())

head(pf.fc_by_age)
tail(pf.fc_by_age)

pf.fc_by_age <- arrange(pf.fc_by_age,age)

head(pf.fc_by_age)
tail(pf.fc_by_age)

#Chaining::

pf.fc_by_age_chain<- pf %>% group_by(age) %>% summarise(fc_mean=mean(friend_count),fc_median=median(friend_count),n=n()) %>% arrange(age)

head(pf.fc_by_age_chain)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age,y=fc_mean),data=pf.fc_by_age) + geom_point()

ggplot(aes(x=age,y=fc_mean),data=pf.fc_by_age) + geom_line(color="blue",size=1)

qplot(age,fc_mean,data=pf.fc_by_age,geom="line")

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age,y=friend_count),data=pf) + 
  geom_point(alpha=1/20,position=position_jitter(h=0),color="orange") + 
  xlim(13,90) + 
  coord_trans(y="sqrt") + 
  geom_line(stat="summary",fun.y=mean) +
  geom_line(stat="summary",fun.y=quantile,probs=0.1,color="blue",linetype=2) +
  geom_line(stat="summary",fun.y=quantile,probs=0.9,color="blue",linetype=2) +
  geom_line(stat="summary",fun.y=quantile,probs=0.5,color="blue")



ggplot(aes(x=age,y=friend_count),data=pf) + 
  coord_cartesian(xlim = c(13, 70),ylim=c(0,1000)) +
  geom_point(alpha=0.05,position=position_jitter(h=0),color="orange") + 
  geom_line(stat="summary",fun.y=mean) +
  geom_line(stat="summary",fun.y=quantile,probs=0.1,color="blue",linetype=2) +
  geom_line(stat="summary",fun.y=quantile,probs=0.9,color="blue",linetype=2) +
  geom_line(stat="summary",fun.y=quantile,probs=0.5,color="blue")


```

#### What are some of your observations of the plot?
Response:
There is similar variation across different summaries mean, median, 10% and 90% quantiles. Variation in the friends count with age is same in each of these summaries.

90% of the users have less than 1000 friends.
50% have less than 400 friends
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor(pf$age,pf$friend_count)

cor.test(pf$age,pf$friend_count,method="pearson")

with(pf,cor.test(age,friend_count,method="pearson"))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.027
***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf,age < 70), cor.test(age, friend_count))

with(subset(pf,age <= 70), cor.test(age, friend_count,method="pearson"))

```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
names(pf)
summary(pf$www_likes_received)
summary(pf$likes_received)
ggplot(aes(x=www_likes_received,y=likes_received),data=pf) + 
  xlim(0,5000) + ylim(0,5000) +
  geom_point(alpha=1/20,position=position_jitter(h=0))

#Checking only the 95 % of the values
ggplot(aes(x=www_likes_received,y=likes_received),data=pf) + 
  xlim(0,quantile(pf$www_likes_received,0.95)) + 
  ylim(0,quantile(pf$likes_received,0.95)) +
  geom_point(alpha=1/20,position=position_jitter(h=0))

#Applying the linear model
ggplot(aes(x=www_likes_received,y=likes_received),data=pf) + 
  xlim(0,quantile(pf$www_likes_received,0.95)) + 
  ylim(0,quantile(pf$likes_received,0.95)) +
  geom_point(alpha=1/20,position=position_jitter(h=0)) +
  geom_smooth(method="lm",color="blue")

with(pf, cor.test(www_likes_received, likes_received))

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
with(pf, cor.test(www_likes_received, likes_received))
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
ggplot(aes(x=www_likes_received,y=likes_received),data=pf) + 
  xlim(0,quantile(pf$www_likes_received,0.95)) + 
  ylim(0,quantile(pf$likes_received,0.95)) +
  geom_point(alpha=1/20,position=position_jitter(h=0)) +
  geom_smooth(method="lm",color="blue")

with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3',dependencies=TRUE)
library(alr3)
library(ggplot2)
data(Mitchell)

names(Mitchell)

head(Mitchell)
tail(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month,y=Temp),data=Mitchell) + geom_point() +
  geom_smooth(method="lm")

ggplot(aes(x=Month,y=Temp),data=Mitchell) + geom_point() +
  geom_smooth()

cor.test(x = Mitchell$Temp,y = Mitchell$Month)
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

#Its around 0.050

cor.test(x = Mitchell$Temp,y = Mitchell$Month)

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=Month,y=Temp),data=Mitchell) + geom_point() + scale_x_discrete(breaks=seq(0,203,12)) + geom_line()

ggplot(aes(x=Month,y=Temp),data=Mitchell) + geom_point() + scale_x_discrete(breaks=seq(0,400,12))


```

***

### A New Perspective

What do you notice?
Response:
Cyclic Nature
Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf.fc_by_age <- arrange(pf.fc_by_age,age)

head(pf)
tail(pf.fc_by_age)

pf$age_with_months <- pf$age + (12- pf$dob_month)/12

pf$age_with_months <- pf$age + (1 - pf$dob_month / 12)

pf$age_with_months <- with(pf, age + (1 - dob_month / 12))


head(pf$age_with_months)
head(pf$age)

```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months<- pf %>% group_by(age_with_months) %>% summarise(friend_count_mean=mean(friend_count),friend_count_median=median(friend_count),n=n()) %>% arrange(age_with_months)


head(pf.fc_by_age_months)


age_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_groups,friend_count_mean = mean(friend_count),friend_count_median = median(friend_count),n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)


```

Programming Assignment
```{r Programming Assignment}
age_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_groups,friend_count_mean = mean(friend_count),friend_count_median = median(friend_count),n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x=age_with_months,y=friend_count_mean),data=subset(pf.fc_by_age_months,age_with_months<71)) + geom_line(color="blue")
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

q1 <- ggplot(aes(x=age_with_months,y=friend_count_mean),data=subset(pf.fc_by_age_months,age_with_months<71)) + geom_line(color="blue") + geom_smooth()

q2 <- ggplot(aes(x=age,y=fc_mean),data=subset(pf.fc_by_age,age<71)) + geom_line(color="blue") + geom_smooth()
q3 <-ggplot(aes(x=round(age/5)*5,y=fc_mean),data=subset(pf.fc_by_age,age<71)) + geom_line(stat="summary",fun.y=mean,color="blue")

library(gridExtra)
grid.arrange(q1,q2,q3)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:
This lesson was very informative about how to analysis two variables using visualization and statistical method. 

It showed the path in which the analysis should be performed. We need to use the data point from the original data-set first and get the feel of the distribution of the data using the variations options like summaries and plots.

Next in order to get better understanding on the data points, we need to look at options of modifying the data points so that we can get more insight of their distribution for different unit of measures.
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

