Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}

pf <- read.delim("pseudo_facebook.tsv")
head(pf,2)


ggplot(aes(x = gender, y = age),data = pf) + 
  geom_boxplot()


ggplot(aes(x = gender, y = age),data = subset(pf,!is.na(gender))) + 
  geom_boxplot()

ggplot(aes(x = gender, y = age),data = subset(pf,!is.na(gender))) + 
  geom_boxplot() +
  stat_summary(fun.y=mean,geom="point",shape=4)


ggplot(aes(x = gender, y = age),data = subset(pf,!is.na(gender))) + 
  geom_boxplot() +
  stat_summary(fun.y=mean,geom="line",shape=4)


pf.fc_by_age_gender <- pf %>% filter(!is.na(gender)) %>% group_by(age,gender) %>% summarise(mean_friend_count = mean(as.numeric(friend_count)), median_friend_count = median(as.numeric(friend_count)), n = n())

head(pf.fc_by_age_gender,2)

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

ggplot(aes(x = age, y = median_friend_count),data = pf.fc_by_age_gender) + geom_line(aes(color=gender))

ggplot(aes(x = age, y = mean_friend_count),data = pf.fc_by_age_gender) + geom_line(aes(color=gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)


head(pf.fc_by_age_gender)
pf.select_age_gender_median <- select(pf.fc_by_age_gender,age,gender,median_friend_count)
pf.reshaped_age_gender_median <- spread(pf.select_age_gender_median,gender,median_friend_count)

pf.fc_by_age_gender.wide <- pf.fc_by_age_gender %>% select(age,gender,median_friend_count) %>% spread(gender,median_friend_count)

head(pf.reshaped_age_gender_median)
head(pf.fc_by_age_gender.wide)


```


***

### Ratio Plot
Notes:

```{r Ratio Plot}


ggplot(data=pf.fc_by_age_gender.wide,aes(x=age,y=female/male)) +
  geom_point()

ggplot(data=pf.fc_by_age_gender.wide,aes(x=age,y=female/male)) +
  geom_line() 


ggplot(data=pf.fc_by_age_gender.wide,aes(x=age,y=female/male)) +
  geom_line() +
  geom_hline(yintercept=1)


ggplot(data=pf.fc_by_age_gender.wide,aes(x=age,y=female/male)) +
  geom_line(color="blue") +
  geom_hline(yintercept=1,linetype=2)


ggplot(data=pf.fc_by_age_gender.wide,aes(x=age,y=female/male)) +
  geom_line(color="blue") +
  geom_hline(yintercept=1,linetype=3)

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}


head(pf$tenure)

pf$year_joined <- floor(2014 - pf$tenure/365)

table(pf$year_joined)

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}


pf$year_joined.bucket <- cut(pf$year_joined,c(2004,2009,2011,2012,2014))


cut(pf$year_joined,breaks=c(2004,2009,2011,2012,2014))

cut(pf$year_joined,breaks=c(2004,2009,2011,2012,2014))

table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}

ggplot(data=subset(pf,!is.na(year_joined.bucket)),aes(x=age,y=friend_count)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=median)



ggplot(data=subset(pf,!is.na(year_joined.bucket)),aes(x=age,y=friend_count)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=mean) 

ggplot(data=subset(pf,!is.na(year_joined.bucket)),aes(x=age,y=friend_count)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=mean) +
  geom_line(stat="summary",fun.y=mean,linetype=2)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}


ggplot(data=subset(pf,!is.na(year_joined.bucket)),aes(x=age,y=friend_count)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=mean) +
  geom_line(stat="summary",fun.y=mean,linetype=2)


```

***

### Friending Rate
Notes:

```{r Friending Rate}

pf_sub <- subset(pf,tenure>0)
pf_sub$rate <- pf_sub$friend_count/pf_sub$tenure
summary(pf_sub$rate)

with(subset(pf,tenure>0),summary(friend_count/tenure))

```

***

### Friendships Initiated
Notes:

What is the median friend rate?
0.2205
What is the maximum friend rate?
417

```{r Friendships Initiated}

head(pf,2)
ggplot(data=subset(pf,tenure>0),aes(x=tenure,y=friendships_initiated/tenure)) +
  geom_line(aes(color=year_joined.bucket))

ggplot(aes(x=tenure,y=friendships_initiated/tenure),data=subset(pf,tenure >=1 )) + 
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=mean)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)



ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
       geom_smooth(aes(color=year_joined.bucket))


```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

Price of the yogurt ranges from 20 to 70.
Majority of them are above 50




```{r Histograms Revisited}

yogurt <- read.csv("yogurt.csv")
names(yogurt)
head(yogurt,2)
tail(yogurt,2)
str(yogurt)
summary(yogurt)

yogurt$id <- factor(yogurt$id)
str(yogurt)


ggplot(data=yogurt,aes(price)) +
  geom_histogram(bin=0.5) +
  ylim(0,200)


ggplot(data=yogurt,aes(price)) +
  geom_histogram(bin=5)


ggplot(data=yogurt,aes(price)) +
  geom_histogram(bin=10)



```

***

### Number of Purchases
Notes:

```{r Number of Purchases}


yogurt <- transform(yogurt,all.purchases = strawberry + blueberry + pina.colada + plain)
names(yogurt)

summary(yogurt$all.purchases)


```

***

### Prices over Time
Notes:

```{r Prices over Time}

ggplot(data=yogurt,aes(x=time,y=price)) +
  geom_jitter(color="blue",alpha=1/20)

ggplot(data=yogurt,aes(x=time,y=price)) +
  geom_jitter(color="blue",alpha=1/10,size=5)


```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

set.seed(4230)
sample.ids <- sample(levels(yogurt$id),16)

ggplot(aes(x=time,y=price),data=subset(yogurt,id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size=all.purchases),pch=1)


```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:
This lesson mainly dealt with the topics 

How to analyse more variables
How to make the visualizations combining multiple variables and make the data talk.
How to form single column using the values from multiple columns
How to reshape the data using the tidyr package in order to spread or gather the data from dataframe.
Generate the heat maps
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

