---
title: "Problem set 4 Revision"
author: "Umesh"
date: "Wednesday, March 04, 2015"
output: html_document
---
Task is to create a scatterplot of price vs x using the ggplot syntax.

```{r}

data(diamonds)

head(diamonds,2)

ggplot(data=diamonds,aes(x=x,y=price)) +
  geom_point()

ggplot(data=diamonds,aes(x=x,y=price)) +
  geom_point(color="blue")

summary(diamonds$price)
summary(diamonds$x)

ggplot(data=diamonds,aes(x=x,y=price,color=cut)) +
  geom_jitter(alpha=1/20) +
  xlim(3,11)

ggplot(data=diamonds,aes(x=x,y=price)) +
  geom_point(color="blue") +
  xlim(3,11)

ggplot(data=diamonds,aes(x=x,y=price,color=cut)) +
  geom_point() +
  xlim(3,11)


```

Observations:

There are few diamonds whose length is 0, these could be wrong entries.
All the diamonds length varies between 4 to 10.74 excluding some having price at 0.
With the increase in the length of the diamonds there is increase in the price of it as well.
Majority if the diamonds are between the range of 4 to 7.5 length



Correlation between price and x:

```{r}

cor.test(x=diamonds$x,y=diamonds$price)
cor.test(x=diamonds$y,y=diamonds$price)
cor.test(x=diamonds$z,y=diamonds$price)

```


Create a simple scatter plot of price vs depth.
```{r}
ggplot(data=diamonds,aes(x=depth,y=price)) +
  geom_point()

ggplot(data=diamonds,aes(x=depth,y=price)) +
  geom_point(color="blue")

ggplot(data=diamonds,aes(x=depth,y=price)) +
  geom_point(alpha=1/100)


ggplot(data=diamonds,aes(x=depth,y=price)) +
  geom_point(alpha=1/100,color="blue") +
  scale_x_continuous(breaks=seq(0,80,2))



ggplot(data=diamonds,aes(x=depth,y=price,color=cut)) +
  geom_point(alpha=1/100) +
  scale_x_continuous(breaks=seq(0,80,2))


```

Correlation between price and depth

```{r}

cor.test(diamonds$price,diamonds$depth)

```

Notes:
There is no correlation between the price and depth of the diamonds.  Correlation co efficient is very very negligible.




Create a scatterplot of price vs carat and omit the top 1% of price and carat values.


```{r}

ggplot(data=diamonds,aes(x=carat,y=price)) +
  geom_point()


ggplot(data=diamonds,aes(x=carat,y=price,color=cut)) +
  geom_point()


ggplot(data=diamonds,aes(x=carat,y=price,color=color)) +
  geom_point()

ggplot(data=diamonds,aes(x=carat,y=price,color=clarity)) +
  geom_point()

```


Create a scatterplot of price vs. volume (x * y * z). This is a very rough approximation for a diamond's volume.


```{r}

diamonds$volume <- diamonds$x * diamonds$y * diamonds$z


summary(diamonds$volume)


ggplot(data=diamonds,aes(x=volume,y=price)) +
  geom_point()

ggplot(data=diamonds,aes(x=volume,y=price)) +
  geom_point() +
  xlim(0,1000)


ggplot(data=diamonds,aes(x=volume,y=price)) +
  geom_point() +
  xlim(0,750)


ggplot(data=diamonds,aes(x=volume,y=price)) +
  geom_jitter(alpha=1/100) +
  xlim(0,750)


```


Observations:

There are few with zero volume.
There is exponential increase in the price of the diamonds as the volume increases.
Almost all the diamonds  volume is less than 500 except few above 500 and less than 1000. Only one at around 3800
Majority of the diamonds are concentrated below the 200 in volume.


Correlation between price and volume

```{r}

cor.test(diamonds$volume,diamonds$price)

with(subset(diamonds,volume>0 & volume <= 800),cor.test(volume,price))

```





```{r}


ggplot(aes(x=volume,y=price),data=diamonds) +
  geom_point(color="blue",alpha=1/30,position=position_jitter(h=0)) +
  geom_smooth(method="lm")


ggplot(aes(x=volume,y=price),data=subset(diamonds,volume>0 & volume <= 800)) +
  geom_point(color="blue",alpha=1/30,position=position_jitter(h=0)) +
  geom_smooth(method="lm")


ggplot(aes(x=volume,y=price),data=subset(diamonds,volume >0 & volume <= 800)) +
  geom_point(color="blue",alpha=1/30,position=position_jitter(h=0)) +
  geom_smooth(method="lm")

```

Use the function dplyr package to create a new data frame containing info on diamonds by clarity.


```{r}

diamondsByClarity <- diamonds %>% group_by(clarity) %>% 
  summary(mean_price = mean(price), median_price = median(price), min_price = min(price), max_price = max(price), n = n())
  
head(diamondsByClarity,2)
```



Your task is to write additional code to create two bar plots on one output image using the grid.arrange() function from the package gridExtra.

```{r}

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))


p1<-ggplot(aes(x=color,y=mean_price),data=diamonds_mp_by_color) + geom_bar(stat="identity")
p2<-ggplot(aes(x=clarity,y=mean_price),data=diamonds_mp_by_clarity) + geom_bar(stat="identity")

grid.arrange(p1,p2)

```

Observation from the above plots:

Mean price of the Diamonds with color J is highest. 
Mean price of Diamonds with color D is lowest. 
Mean price of Diamonds with SI2 clarity is highest
Mean price of Diamonds with VVS1 clarity is lowest.
And there is no much difference in mean prices between other colored or clarity diamonds.



