---
title: "ProblemSet2_V1"
author: "Umesh"
date: "Tuesday, February 17, 2015"
output: html_document
---


Load the diamonds dataset and check the dataset

```{r}
library(ggplot2)
data(diamonds)

head(diamonds)
tail(diamonds)

```


Scatterplot of price vs x

```{r}
ggplot(aes(x=x,y=price),data=diamonds) + geom_point(color="blue")

ggplot(aes(x=x,y=price),data=diamonds) + geom_point(color="blue") + scale_x_continuous(limit=c(3,10),breaks=c(0,11,1))


ggplot(aes(x=x,y=price),data=diamonds) + geom_point(color="blue") + scale_x_continuous(limit=c(3,11))

ggplot(aes(x=x,y=price),data=diamonds) + geom_point(alpha=1/20,color="blue") + scale_x_continuous(limit=c(3,11))

ggplot(aes(x=x,y=price),data=diamonds) + geom_point(color="blue") + scale_x_continuous(limit=c(3,11))

ggplot(aes(x=x,y=price),data=diamonds) + geom_point(color="blue") + scale_x_continuous(limit=c(3,11)) + geom_smooth(color="red")

summary(diamonds)

cor.test(diamonds$price,diamonds$x)
cor.test(diamonds$price,diamonds$y)
cor.test(diamonds$price,diamonds$z)
cor(diamonds$price,diamonds$x)
```

Observations about the above plot?

There are few diamonds whose length is 0, these could be wrong entries.
All the diamonds length varies between 4 to 10.74 excluding some having price at 0.
With the increase in the length of the diamonds there is increase in the price of it as well


Create a simple scatter plot of price vs depth.
```{r}
ggplot(aes(x=price),data=diamonds) + geom_histogram()
ggplot(aes(x=depth),data=diamonds) + geom_histogram()

ggplot(aes(x=depth,y=price),data=diamonds) + geom_point(color="blue")
```

Change the code to make the transparency of the points to be 1/100 of what they are now and mark the x-axis every 2 units
```{r}
ggplot(aes(x=depth,y=price),data=diamonds) +
  geom_point(color="blue",alpha=1/100) +
  scale_x_continuous(breaks=seq(0,80,2))
```

Correlation?
```{r}
cor.test(diamonds$depth,diamonds$price)
```


Create a scatterplot of price vs carat and omit the top 1% of price and carat values.

```{r}
ggplot(aes(x=carat,y=price),data=diamonds) +
  geom_point(color="blue")
```

Create a scatterplot of price vs. volume (x * y * z).This is a very rough approximation for a diamond's volume.
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
ggplot(aes(x=volume,y=price),data=diamonds) +
  geom_point(color="blue")

# count(diamonds$volume == 0)
# detach("package:plyr", unload=TRUE)

data1 <- subset(diamonds,volume!=0 & volume <800)
head(data1)
cor.test(data1$price,data1$volume)

```

Subset the data to exclude diamonds with a volume
greater than or equal to 800. Also, exclude diamonds
with a volume of 0. Adjust the transparency of the
points and add a linear model to the plot. (See the
Instructor Notes or look up the documentation of
geom_smooth() for more details about smoothers.)

```{r}
diamonds <- subset(diamonds,volume!=0 & volume <800)

ggplot(aes(x=volume,y=price),data=diamonds) +
  geom_point(color="blue",alpha=1/30,position=position_jitter(h=0)) +
  geom_smooth(method="lm")

```

Use the function dplyr package
to create a new data frame containing
info on diamonds by clarity.

Name the data frame diamondsByClarity

The data frame should contain the following
variables in this order.
(1) mean_price
(2) median_price
(3) min_price
(4) max_price
(5) n

```{r}
library(dplyr)

diamond_clarity <- group_by(diamonds,clarity)
diamondsByClarity <- summarise(diamond_clarity,mean_price = mean(price),median_price=median(price), min_price=min(price), max_price=max(price), n = n())

head(diamondsByClarity)

```


```{r}
data(diamonds)
library(dplyr)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

head(diamonds_mp_by_clarity)
p1<-ggplot(aes(x=color,y=mean_price),data=diamonds_mp_by_color) + geom_bar(stat="identity")
p2<-ggplot(aes(x=clarity,y=mean_price),data=diamonds_mp_by_clarity) + geom_bar(stat="identity")

library(gridExtra)

grid.arrange(p1,p2)
```

