---
title: "Problemset 5"
author: "Umesh"
date: "Friday, March 06, 2015"
output: html_document
---

Create a histogram of diamond prices.Facet the histogram by diamond color and use cut to color the histogram bars.


```{r}

head(diamonds)

ggplot(data=diamonds,aes(price)) +
  geom_histogram(aes(color=cut)) + 
  scale_x_log10() +
  facet_wrap(~color) 



ggplot(data=diamonds,aes(price,fill=cut)) +
  geom_histogram(aes(color=cut)) + 
  scale_x_log10() +
  facet_wrap(~color) 


ggplot(data=diamonds,aes(price,fill=cut)) +
  geom_histogram() + 
  scale_x_log10() +
  facet_wrap(~color) 


```



# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

```{r}


ggplot(data=diamonds,aes(x=table,y=price,color=cut)) +
  geom_point() +
  scale_x_continuous(limits=c(50,80),breaks=seq(50,80,2))


ggplot(data=diamonds,aes(x=table,y=price,color=cut)) +
  geom_point(size=5) +
  scale_x_continuous(limits=c(50,80),breaks=seq(50,80,2))


by(diamonds$table,diamonds$cut,summary)

?summary
```



# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.


```{r}

ggplot(data=subset(diamonds,x*y*z>0),aes(x=x*y*z,y=price)) +
  geom_point() +
  scale_y_log10() +
  xlim(0,400)


ggplot(data=subset(diamonds,x*y*z>0),aes(x=x*y*z,y=price,color=clarity)) +
  geom_point(size=4.5) +
  scale_y_log10() +
  xlim(0,400)



```

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

```{r}
pf <- read.delim("pseudo_facebook.tsv")

head(pf,2)
tail(pf,2)


pf$prop_initiated <- pf$friendships_initiated/pf$friend_count
head(pf$prop_initiated)
tail(pf$prop_initiated)


```




# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.


```{r}

head(pf)
pf$year_joined <- floor(2014-pf$tenure/365)

head(pf$year_joined)

head(!is.nan(pf$prop_initiated))

pf$year_joined.bucket <- cut(pf$year_joined,breaks=c(2004,2009,2011,2012,2014))

head(pf$year_joined.bucket)

ggplot(data=pf,aes(x=tenure,y=prop_initiated)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=median)

ggplot(data=subset(pf,!is.nan(prop_initiated)),aes(x=tenure,y=prop_initiated)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=median)


ggplot(data=subset(pf,!is.nan(prop_initiated)),aes(x=tenure,y=prop_initiated)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=median) +
  geom_smooth()


ggplot(data=subset(pf,!is.nan(prop_initiated)),aes(x=tenure,y=prop_initiated)) +
  geom_line(aes(color=year_joined.bucket),stat="summary",fun.y=median) +
  geom_smooth(method="lm")

pf <- subset(pf,!is.nan(prop_initiated))

by(pf$prop_initiated,pf$year_joined.bucket,mean)

```


# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.


```{r}


ggplot(data=diamonds,aes(x=cut,y=price/carat)) +
  geom_jitter(aes(color=color)) +
  facet_wrap(~clarity)


```

