Read and get the feel of the Diamond dataset 

```{r}
data(diamonds)
dim(diamonds)
str(diamonds)
summary(diamonds)

```

Create the histogram of the price of all the diamonds in the diamonds data set

```{r}

ggplot(aes(price),data=diamonds) +
  geom_histogram()

ggplot(aes(price),data=diamonds) +
  geom_histogram(binwidth=200)


```


Check the summary of the price of the diamonds in the diamonds dataset

```{r}
summary(diamonds$price)
```

Notes on the summary of the price of the diamonds:

Shape is right skewed with the center of the distribution being at 2401.

Since the data is skewed median is the best measure of center.

Median  = 2401
Mean = 3933
Min = 326
Max = 18820


Check the Diamonds count based on the price:

```{r}

head(diamonds,2)

count(diamonds,price<500)
count(diamonds,price<250)
count(diamonds,price>=15000)

```


Explore the largest peak in the price histogram you created earlier.

Try limiting the x-axis, altering the bin width,and setting different breaks on the x-axis.


```{r}

ggplot(aes(price),data=diamonds) + geom_histogram(color="black",fill="blue")


ggplot(aes(price),data=diamonds) + geom_histogram(color="black",fill="blue",binwidth=100)


ggplot(aes(price),data=diamonds) + 
  geom_histogram(color="black",fill="blue",binwidth=50) +
  scale_x_continuous(limits=c(0,5000))


ggplot(aes(price),data=diamonds) + 
  geom_histogram(color="black",fill="blue",binwidth=25) +
  scale_x_continuous(limits=c(0,5000),breaks=seq(0,5000,250)) 


```

Notes on the distribution:

There are no diamonds with the price $1500.
THere are more diamonds within the price range of $500 to $750
There is no much difference in the distribution of the diamonds with price range of $2500 to $5000


Break out the histogram of diamond prices by cut.

```{r}
ggplot(aes(price),data=diamonds) + 
  geom_histogram(color="black",fill="blue",binwidth=25) +
  scale_x_continuous(limits=c(0,5000),breaks=seq(0,5000,250)) +
  facet_wrap(~cut)


ggplot(aes(price),data=diamonds) + 
  geom_histogram(fill="maroon",binwidth=25) +
  scale_x_continuous(limits=c(0,5000),breaks=seq(0,5000,250)) +
  facet_wrap(~cut)


ggplot(aes(price),data=diamonds) + 
  geom_histogram(fill="purple",binwidth=25) +
  scale_x_continuous() +
  facet_wrap(~cut)

```

Notes :
The distribution of the diamonds based on the price among different cuts is not the same. 
There are more diamonds with ideal and very good cut
There are very few diamonds in the fair cut.


Summary of price by cut:

```{r}

by(diamonds$price,diamonds$cut,summary)
```


Make the y axis free for the facet wrap 


```{r}

ggplot(aes(price),data=diamonds) + 
  geom_histogram(fill="maroon",binwidth=25) +
  scale_x_continuous(limits=c(0,5000),breaks=seq(0,5000,250)) +
  facet_wrap(~cut,scales="free_y")

```

Create a histogram of price per carat and facet it by cut. You can make adjustments
to the code from the previous exercise to get started.

```{r}

ggplot(aes(price/carat),data=diamonds) + 
  geom_histogram(fill="maroon",binwidth=25) +
  facet_wrap(~cut,scales="free_y")


ggplot(aes(price/carat),data=diamonds) + 
  geom_histogram(fill="maroon",binwidth=0.005) + 
  scale_x_log10() +
  facet_wrap(~cut,scales="free_y")


ggplot(aes(log10(price/carat)),data=diamonds) + 
  geom_histogram(fill="maroon",binwidth=0.025) + 
  facet_wrap(~cut,scales="free_y")

```


Notes :
Distribution of the price per carat is normally distributed across different cuts except in Very Good cut where the distribution is slightly bimodal



Investigate the price of diamonds using box plots,numerical summaries, and one of the following categorical variables: cut, clarity, or color.


```{r}

ggplot(aes(x=cut,y=price,color=cut),data=diamonds) +
  geom_boxplot() + coord_cartesian(ylim=c(0,7000))

ggsave("box_plot_pricebycut.jpg",dpi=300,width=10,height=5)

ggplot(aes(x=color,y=price,color=color),data=diamonds) +
  geom_boxplot() + coord_cartesian(ylim=c(0,10000))

ggsave("box_plot_pricebycolor.jpg",dpi=300,width=10,height=5)

ggplot(aes(x=clarity,y=price,color=clarity),data=diamonds) +
  geom_boxplot() +coord_cartesian(ylim=c(0,7500))

ggsave("box_plot_pricebyclarity.jpg",dpi=300,width=10,height=5)

by(diamonds$price,diamonds$cut,summary)

by(diamonds$price,diamonds$clarity,summary)

by(diamonds$price,diamonds$color,summary)

b1 <- qplot(x = cut, y= price, data = diamonds,geom='boxplot',color=cut) 

b2 <- qplot(x = cut, y= price, data = diamonds,geom='boxplot',color=cut) + coord_cartesian(ylim = c(0, 7000))

grid.arrange(b1,b2,ncol=1)


```



Investigate the price per carat of diamonds across the different colors of diamonds using boxplots.

```{r}

ggplot(aes(x=color,y=price/carat,color=color),data=diamonds) +
  geom_boxplot() +coord_cartesian(ylim=c(0,7500))

ggsave("box_plot_pricepercarat_colorwise.jpg",dpi=300,width=10,height=5)


b1 <-  qplot(x = color, y= price/carat, data = diamonds,geom='boxplot',color=color) 

b2 <- qplot(x = color, y= price/carat, data = diamonds,geom='boxplot',color=color) + coord_cartesian(ylim = c(0, 6000))

grid.arrange(b1,b2,ncol=1)

by(diamonds$price/diamonds$carat,diamonds$color,summary)

```


Frequency polygon for carat

```{r}

ggplot(aes(carat),data=diamonds) +
  geom_freqpoly(color="blue")



ggplot(aes(carat),data=diamonds) +
  geom_freqpoly(color="blue",binwidth=0.05)

ggplot(aes(carat),data=diamonds) +
  geom_freqpoly(color="blue",binwidth=0.025) +
  scale_x_continuous(breaks=seq(0,5,0.05))
  

```


